Easy as Pi, Volume II: On the Goodput and Energy of Network Coding Schemes in the Raspberry Pi

### 2016-06-28 introduction.tex
Abstract changes

% RAW INTRODUCTION: NEED OF DISTRIBUTED NETWORKS AND THE RASPBERRY PI
Due to the advent of the \ac{IoT}, around 50 million devices
ranging from sensors to phones are expected to be connected into
data networks in relatively short period of time \textbf{[REFERENCE]}.
This massive deployment requires the design and testing of new
distributed systems that permit to cope with the amount of traffic
from the proposed services provided by these devices. Therefore,
development platforms that help to quickly deploy, analyze and
evaluate this type of settings are highly desirable for research.
The emergence of a powerful and inexpensive portable computer such
as the Raspberry Pi running standard \ac{OS} like Linux or even
Windows allows to achieve this goal since enables the deployment
of distributed systems at a lower cost than the necessary in
previous years by using standard protocol. We also complement this
with an in-depth evaluation of the Raspberry Pi as a platform to
enable a new paradigm in networking and storage known as \ac{NC},
which relies on processing at intermediate nodes to achieve
higher throughput, delay and reliability performance than
state-of-the-art routing techniques.

% INTRODUCTION TO NETWORK CODING
Introduced by Ahlswede et al.~\cite{ahlswede2000network}, network
coding constitutes a paradigm shift in the way how researchers and
industry understand and operate networks, by changing the role of
intermediate relays in the process of transmission of information.
Relays are no longer limited to storing and forwarding data, but also
take part in the coding process, through a process called recoding,
where the relay generates new linear combinations of incoming coded
packets without previously decoding the data. Network coding allows
the increase of throughput, reliability, security and delay
performance of the networks.

Compared to other broadly used coding schemes, e.g., Gallager's LDPC
codes~\cite{gallager1962low} or Reed-Solomon
codes~\cite{reed1960polynomial}, network coding is a technology that
has been implemented in real systems since the early years of its
conception. In 2006, Katabi et al.~\cite{katabi2006practical} published
the results of the performance of their protocol COPE when implemented
in a real wireless system, which relied on minimalistic coding to
provide interesting gains. In 2008, Pedersen et
al.~\cite{pedersen2008implementation} used commercially available
Symbian OS enabled mobile phones to implement network coding in a
\ac{D2D} cooperation-based application. In 2011, researchers at Aalborg
University developed Kodo~\cite{kodo2011pedersen}, a C++11 network
coding library with open source code for researchers intended to make
network coding implementations easily available for both, the research
community and commercial entities.
Another case is CATWOMAN~\cite{hundeboll2012catwoman}, a protocol
implemented on top of the BATMAN protocol~\cite{johnson2008simple} for
IEEE 802.11 multi-hop meshed networks using some of the intuition from
COPE, but deploying it in real systems. It is currently available as
open source code in the Linux kernel. Many other implementations have
been tested on real world systems~
\cite{pahlevani2013playncool,katti2008xors,krigslund2013core,paramanathan2013leanandmean}.

Despite this successful deployment in real systems, many of these
protocols and contributions have been implemented in separate testbeds
and the experiences are hard to reproduce.  Given the need for
reproducibility of past and future results among researchers and the
need of low-cost and easy-to-deploy testbeds, this paper contribution
is provide detailed instructions on how to set up a low-cost,
heterogeneous and scalable testbed consisting of Raspberry Pis
(Raspis), and by providing detailed measurements of goodput and energy
consumption of such testbed when performing network coding operations
with different codecs (such as full \ac{RLNC} \cite{ho2006random},
multicore-enabled \ac{RLNC}, sparse \ac{RLNC} and tunable sparse \ac{RLNC}).
The heterogeneity of the testbed comes from the fact that old, single core,
Raspberry Pi 1 model B units are integrated seamlessly in the described
testbed with newer, multicore versions, like the Raspberry Pi 2 model B.
Our testbed allows us to obtain a comprehensive set of measurements
to characterize the goodput and energetic footprint of these devices
showing that processing speeds of 800 Mbps and energy per bit values
of 1 nJ or even one order of magnitude less are possible.

Our work is organized as follows. Section~\ref{sec:schemes} defines the
coding schemes employed in our study. Later, in Section~\ref{sec:metrics}
we describe the considered metrics and methodology for performance comparison
of the codes deployed in the \ac{Raspi}. In Section~\ref{sec:measurements}, we
show the measurements in the \ac{Raspi} models of the mentioned metrics
providing full discussions about the observed operational regimes. Final
conclusions and future work are reviewed in Section~\ref{sec:conclusions}.

>>>>

% RAW INTRODUCTION: NEED OF DISTRIBUTED NETWORKS AND THE RASPBERRY PI
Due to the advent of the \ac{IoT}, approximately fifty million devices
ranging from sensors to phones are expected to be connected through
data networks in relatively short period of time \textbf{[REFERENCE]}.
This massive deployment requires the design and testing of new
distributed systems that permit to cope with the amount of traffic
from the proposed services provided by these devices. Therefore,
development platforms that help to quickly deploy, analyze and
evaluate this type of settings are highly desirable for research.
The emergence of a powerful and inexpensive portable computer such
as the Raspberry Pi, running standard \ac{OS} like Linux or Windows,
allows to achieve this goal since it enables the deployment
of distributed systems at a low cost. Also, the use of an \ac{OS} that
has a large user community helps to manage this networks in less time
\textbf{INCLUDE PRIOR WORK}.

% INTRODUCTION TO NETWORK CODING
Introduced in~\cite{ahlswede2000network}, \ac{NC}
constitutes a paradigm shift in the way networks are understood
by changing how information is sent through them.
Instead of treating the packets as atomic
unmodifiable units, packets are seen as algebraic entities in a \ac{GF}
that can be operated on to create new coded packets. This permits to
remove the limitation of sending \textit{specific} packets by now sending
coded packets as \textit{linear equations} of the original ones. This
change in the way of seeing how the data is represented, brings new
features that can be exploited. In this way, instead of typically encoding
and decoding on a hop basis, relaying nodes can take part in the
coding process without needing to decode. Therefore, a relay can
\textit{recode} packets, i.e. encode again previously received encoded
(but not decoded) packets in order to reduce delay and still
take advantage of the data representation for the next hop.

Compared to other broadly used coding schemes, e.g. \ac{LDPC}
codes~\cite{gallager1962low} or Reed-Solomon
codes~\cite{reed1960polynomial}, network coding is a technology that
has been implemented in real systems since the early years of its
conception. Katti et al.~\cite{katti2008xors} showed the
performance of the COPE protocol for the two-way relay channel
in a wireless network, which relied on minimalistic coding and obtaining
gains over a forwarding scheme. Later, Pedersen et
al.~\cite{pedersen2008implementation} used commercially available
Symbian OS mobile phones to implement network coding in a
\ac{D2D} cooperation-based application. Furthermore, researchers at Aalborg
University developed Kodo~\cite{kodo2011pedersen}, a C++11 network
coding library intended to make network coding functionalities easily
available for both the research community and commercial entities.
Based on Kodo is CATWOMAN~\cite{hundeboll2012catwoman}, a protocol
implemented on top of the BATMAN protocol~\cite{johnson2008simple} for
WiFi multi-hop meshed networks using some of the intuition from
COPE, but deployed within the X topology with overhearing and
currently available as open source in the Linux kernel. Many other
implementations have been tested on real world systems~
\cite{pahlevani2013playncool,katti2008xors,krigslund2013core,paramanathan2013leanandmean}.

In this work, we provide detailed measurements of the goodput and energy
consumption of \ac{Raspi} models 1 and 2, when performing network coding
operations with different codecs based on \ac{RLNC} \cite{ho2006random}
such as: full \ac{RLNC}, multicore-enabled \ac{RLNC}, sparse \ac{RLNC}
and tunable sparse \ac{RLNC}. For this work, we assess the \ac{Raspi}
performance excluding the channel effect in the processing, meaning that
we do not consider packet losses or delays in order to have a
description of the codes in terms of their parameters. To achieve this,
we perform a measurement campaign with the indicated coding schemes and
their parameters in our local Raspberry Pi testbed. Our measurements permit
us to characterize the processing speed (goodput) and energy consumption
of these devices showing that speeds of 800 Mbps and processing energy per bit
values of 0.1 nJ are possible.

Our work is organized as follows. Section~\ref{sec:schemes} defines the
coding schemes employed in our study. Later, in Section~\ref{sec:metrics}
we describe the considered metrics and methodology for performance comparison
of the codes deployed in the \ac{Raspi}. In Section~\ref{sec:measurements}, we
show the measurements in the \ac{Raspi} models of the mentioned metrics
providing full discussions about the observed operational regimes and effects.
Final conclusions and future work are reviewed in
Section~\ref{sec:conclusions}.
